---
title:  "Statistical Comparision of NASA-TROPOMI - Satellite and GEOS-CF - Model Data"
author: Shruti Jagini 
---

# Introduction
Wildfires in California occur regularly, these have significantly contributed in shaping the nature over the past century. Southern California is the highest fire prone regions of the world, its predominant forests have evolved to thrive on the phenomenon of wildfires. Over the past decades the fires have caused significant damage and loss of both property, human and animal life. As a result of these wildfires, a number of pollutants such as, oxides of nitrogen, oxides of sulfur, methane, carbon monoxide are released into the atmosphere. This project deals with statistical comparison between the satellite data (NASA-TROPOMI), and a geographical model data (GEOS-CF) with respect to NO2 and CO concentrations emitted from California wildfires.

#### NASA-TROPOMI 
The TROPOspheric Monitoring Instrument (TROPOMI) is a satellite instrument which is on board the Copernicus Sentinel-5 Precursor satellite. The Sentinel-5 Precursor (S5P) is launched by the European Space
Agency (ESA). The tropospheric monitoring instrumentcovers the wavelength of ultraviolet-visible (270 nm to 495 nm), near infrared region (675 nm to 775 nm) and shortwave infrared (2305 nm - 2385 nm).This instrument provides measurements on concentrations of various pollutants and components of atmosphere like ozone, NO 2 , SO 2 ,
CH 4 , CO, formaldehyde, aerosols and clouds.

#### GEOS-CF
The GEOS-Chem is a global 3-D model which is used for monitoring atmospheric chemistry, it is driven by assimilated meteorological observations from the Goddard Earth Observing System (GEOS).This model takes into account about 220 reactive species and 720 reactions. GEOS-CF works by taking into account atmospheric dynamics such as wind speed and chemistry of pollutants (such as their chemical reactions, dispersion, reactivity) in predicting their concentrations

# Materials and methods

[~ 200 words]

Narrative: Clear narrative description of the data sources and methods. Includes data from at least two sources that were integrated / merged in R.

Code: The code associated with the project is well organized and easy to follow. Demonstrates mastery of R graphics and functions.

Data: The underlying data are publicly accessible via the web and downloaded/accessed within the Rmd script. If you want to use your own data, you must make it available on a website (e.g. Figshare) so that others are able to re-run your code.

You can do bullets like this:

* The first most important thing
* The second most important thing
* The third most important thing

You can do numbers like this:

1. The first most important thing
2. The second most important thing
3. The third most important thing

See [http://rmarkdown.rstudio.com/](http://rmarkdown.rstudio.com/) for all the amazing things you can do.


Here's my first code chunk.
```{r}
1+2
```

Load any required packages in a code chunk (you may need to install some packages):

```{r, message=F, warning=F}
library(tidyverse)
library(leaflet)
library(kableExtra)
knitr::opts_chunk$set(cache=TRUE)  # cache the results for quick compiling
```

## Download and clean all required data
```{r}
n=20
data=data.frame(x=runif(n,-180,180),
                y=runif(n,-60,60),
                size = runif(n, 5, 20),
                category = factor(
                  sample(letters[1:5], n, replace = TRUE)
                  ),
                value = rnorm(n))
```

```{r, results='asis'}
data %>% 
  slice(1:10) %>% #show only 1:n rows
  kable(digits=2,align="c")%>% #make table and round to two digits
  kable_styling(bootstrap_options = 
                  c("striped", "hover", "condensed", "responsive")) #apply other formatting
```

Add any additional processing steps here.

# Results

[~200 words]

Tables and figures (maps and other graphics) are carefully planned to convey the results of your analysis. Intense exploration and evidence of many trials and failures. The author looked at the data in many different ways before coming to the final presentation of the data.

Show tables, plots, etc. and describe them.

```{r, fig.width=6, fig.height=3, fig.cap="Map of completely random data"}
m <- leaflet(data) %>% 
  addTiles() %>% 
  addCircleMarkers(~x, ~y, radius = ~size,color = ~as.factor(category)) %>% 
  addPopups(~x[2], ~y[2], "Random popup")
m  # a map with the default OSM tile layer
```


```{r}
data %>% 
  ggplot(aes(x=x,y=y,col=category))+
  geom_point()
```

# Conclusions

[~200 words]

Clear summary adequately describing the results and putting them in context. Discussion of further questions and ways to continue investigation.

# References

All sources are cited in a consistent manner
